---
title: "Titanic survivors"
author: "Michel Mariën"
date: "21 november 2018"
output: 
  ioslides_presentation: 
    fig_caption: yes
    keep_md: yes
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(ggplot2)
library(ggrepel)

```

## Surviving the Titanic

In this document we're going to explore the question whether women had preferential treatment during
the evacuation of Titanic and thus a higher chance of survival

```{r survival,fig.align="center",out.width="100%"}

Titanic_surv<- as.data.frame(Titanic) %>% group_by(Survived) %>% 
  summarise(Surv_sum=sum(Freq))%>% 
  mutate(Surv=Surv_sum/sum(Surv_sum))

colors <- c("red", "green")

plot_ly(Titanic_surv, labels = ~Survived, values = ~Surv, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',
        text = ~paste(round(Surv*100,2), "%"),
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
                      #The 'pull' attribute can also be used to create space between the sectors
        showlegend = FALSE) %>%
  layout(title = 'Survival rate Titanic',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```

## Data exploration

We are going to look at several factors that might shed some light on chance of survival

- Sex
- Age
- Class

## Survival chances based on Sex

```{r Sex,fig.align="center",out.width="100%"}

Titanic_surv_sex<- as.data.frame(Titanic) %>% group_by(Sex,Survived) %>% 
  summarise(Surv_sum=sum(Freq))%>% 
  mutate(Sex_surv=Surv_sum/sum(Surv_sum))

splitted_list <- split(Titanic_surv_sex, Titanic_surv_sex$Sex)

plot_list <- lapply(splitted_list, plot_ly, x = ~Survived, y = ~Sex_surv, type = "bar", text = ~paste(Sex,round(Sex_surv*100,2), '%'),hoverinfo = 'text',showlegend = FALSE)

plot_list2 <- lapply(plot_list,layout,title="Bar chart with survival rates for Males (left) and Females (right)")


subplot(plot_list2,nrows=1, margin = .05)
```
 
<font size="4"> One can conclude from the chart above that being a woman significantly improved 
your chances of survival </font>
 
## Survival chances based on Age

```{r age,fig.align="center",out.width="100%",warning=FALSE}

Titanic_surv_sex<- as.data.frame(Titanic) %>% group_by(Sex,Age,Survived) %>% 
  summarise(Surv_sum=sum(Freq))%>% 
  mutate(Sex_surv=Surv_sum/sum(Surv_sum))

#colors <- c("red", "green","blue")
#,showlegend = FALSE,marker = list(colors = colors)
splitted_list <- split(Titanic_surv_sex, Titanic_surv_sex$Sex)

plot_list <- lapply(splitted_list, plot_ly, x = ~Age, y = ~Sex_surv,color=~Survived, type = "bar", text = ~paste(Survived,round(Sex_surv*100,1), '%'),hoverinfo = 'text',showlegend = FALSE)

#plot_list2 <- lapply(plot_list,add_annotations,xref = 'x', yref = 'y',x = ~Age, y = ~Sex_surv,text = ~Survived,showarrow = FALSE)

plot_list2 <- lapply(plot_list,layout,title="Bar chart with survival rates for Males (left) and Females (right)", barmode = 'stack')

subplot(plot_list2,nrows=1, margin = .05)

```

<font size="4"> Being a child improves your chances of survival as a male but being a child and female does not improve your chances in comparison to adult females </font>

 
## Survival chances based on Class
 
```{r class,fig.align="center",out.width="100%",warning=FALSE}

Titanic_surv_sex<- as.data.frame(Titanic) %>% group_by(Sex,Class,Age,Survived) %>% 
  summarise(Surv_sum=sum(Freq))%>% 
  mutate(Sex_surv=Surv_sum/sum(Surv_sum))

splitted_list <- split(Titanic_surv_sex, list(Titanic_surv_sex$Sex,Titanic_surv_sex$Age))

plot_list <- lapply(splitted_list, plot_ly, x = ~Class, y = ~Sex_surv,color=~Survived, type = "bar", text = ~paste(Sex,Age,Survived,round(Sex_surv*100,1), '%'),hoverinfo = 'text',showlegend = FALSE)

plot_list2 <- lapply(plot_list,layout,title="Bar chart with survival rates per class for Children (left) and Adults (right)", barmode = 'stack')

subplot(plot_list2,nrows=1, margin = .05)

```

<font size="3"> Apparently, if you traveled in first of second class, you had the highest change of survival. Being a child in those classes (males and females) gave you a 100% change of living another day. It seems that passengers in the higher classes got a preferential treatment not females  </font>
